At the end of this worksheet you should be able to:
\begin{itemize}
	\item convert units within the metric system.
	\item apply how to correctly convert area and volume units.
	\item apply scientific notation.
	\item apply the definition of percent change.
	\item convert between a ratio statement and percent change statement.
	\item apply the two forms of proportionality statements to make predictions from dependent variable change. 
\end{itemize}


\begin{enumerate}
\setlength\itemsep{1 in}

\item
How many meters are in a decameter and how do you know?\\
	
\item
How many decameters are in a meter and why?\\

\item 
How many centimeters are in a meter? Draw on your paper approximately a centimeter.\\

\item 
How many millimeters are in a meter and how many meters are in a millimeter?\\

\item 
The diameter of the earth is 6,380,000 m. What is this in kilometers?\\

\item 
How many centimeters are in 3.2 km?\\

\item 
How many inches are in 1 meter? Do this using the conversion of 1 in = 2.54 cm.\\

\item 
If I say your desk has an area of 0.5 meters, then what is wrong with that statement?\\

\item 
Convert miles per hour to meters per second.\\
\\


\item
List as many formulas for the area of different shapes as you can remember (or look some up). What do these all have in common? What about volume formulas? What does this tell you about the units of these kinds of quantities?\\
\\


\item 
If your desk has an area of \SI{0.5}{meters^2}, then what is its area in \si{centimeters^2}?\\

\item 
If a ball has a diameter of 18 cm, then what is the volume of the ball in \si{meters^3}?\\

\item Calculate your age in seconds on your last birthday.

\item 
Make up your own unit conversion problem that involves at least two conversions chained together.\\
\\

\item 
Put the number 21,345,000,000 kg in scientific notation?\\

\item 
Put the number 0.0000000234 km in scientific notation?\\

\item 
Correct the scientific notation of $140\times10^{-3}$ seconds.

\item 
Correct the scientific notation of $0.012\times10^{-3}$ meters.

\item 
What is \SI{0.000345}{meter} in micrometers \si{\um}?

\item 
What is \SI{3.4e5}{\meter} in kilometers?

\item 
If I have three times as many marbles as you do then what is the ratio of my marbles to yours? What is the ratio of your marbles to mine?


\item If I increase in speed from 25 mph to 40 mph, then by what percent has my speed changed? By what ratio has the speed changed?\\

\item If my speed changes from 29 mph to 10 mph, then by what percent has my speed changed? What is the negative sign in the answer tell you? By what factor has your speed changed from initially to finally? Could a negative ratio make sense here?\\

\item If the generic variable $y$ is inversely proportional to the a variable $x$, then write out this statement mathematically in two ways.\\

\item Since the area of a triangle is directly proportional to both the base and height, then how would an \emph{equation} for this look like and what is the constant of proportionality?\\

\item The formula for the volume of a cylinder is $V=\pi r^2 h$. How would you write out a proportionality statement that was consistent with this formula? What is the constant of proportionality?\\

\item If the radius of a sphere changes by a factor of 2.7, then by what factor does the volume change?\\

\item How would you turn the previous problem "inside out"? 

\item By what percent does the volume change in the previous two problems?
	
\end{enumerate}